---
title: "analysis"
---

First the packages are loaded:

```{r}
#| output: false
library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)
```

The first ten lines of the babynames dataset look like:

```{r}
head(babynames) |> kable()
```

Functions to get most frequent babynames over a time period:

```{r}
#| code-fold: true
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

Plotting the most frequent babynames for girls from 1950 onwards:

```{r}
#| label: fig-line-girls
#| fig-cap: "Line plot displaying proportion of top 10 girl names by year"
get_most_frequent(babynames, select_sex = "F") |>
  plot_top()
```

In @fig-line-girls you can see that the name 'Lisa' peaked at the end of the 60s, followed by two major peaks of 'Jennifer' and 'Jessica'.  

Plotting the most frequent babynames for girls from 1950 onwards:

```{r}
#| label: fig-line-boys
#| fig-cap: "Line plot displaying proportion of top 10 boy names by year"
get_most_frequent(babynames, select_sex = "M") |>
  plot_top()
```

@fig-line-boys shows that in general the ten most popular boys names became a lot less popular after the 1990s.
