---
title: "analysis"
---

First the packages are loaded:

```{r}
#| output: false
library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)
```

The first ten lines of the babynames dataset look like:

```{r}
head(babynames) |> kable()
```

Functions to get most frequent babynames over a time period:

```{r}
#| code-fold: true
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

Plotting the most frequent babynames for girls from 1950 onwards:

```{r}
#| label: fig-line-girls
#| fig-cap: "Line plot displaying proportion of top 10 girl names by year"
get_most_frequent(babynames, select_sex = "F") |>
  plot_top()
```

In @fig-line-girls you can see that the name 'Lisa' peaked at the end of the 60s, followed by two major peaks of 'Jennifer' and 'Jessica'.

Plotting the most frequent babynames for girls from 1950 onwards:

```{r}
#| label: fig-line-boys
#| fig-cap: "Line plot displaying proportion of top 10 boy names by year"
get_most_frequent(babynames, select_sex = "M") |>
  plot_top()
```

@fig-line-boys shows that in general the ten most popular boys names became a lot less popular after the 1990s.

```{r}
#| label: fig-line-2010
#| layout: [[50,50], [100]]
#| fig-cap: "Most popular girl names from 2010 onwards"
#| fig-subcap: 
#|   - "Top 5"
#|   - "Top 10"
#|   - "Top 30"
# get most frequent girl names from 2010 onwards
from_year <- 2010
most_freq_girls <- get_most_frequent(babynames, select_sex = "F",
                                     from = from_year)

# plot top 5 girl names
most_freq_girls |>
  plot_top(top = 5)

# plot top 10 girl names
most_freq_girls |>
  plot_top(top = 10)

# get top 30 girl names in a matrix
# with names in rows and years in columns
prop_df <- babynames |> 
  filter(name %in% most_freq_girls$most_frequent$name[1:30] & sex == "F") |>
  filter(year >= from_year) |> 
  select(year, name, prop) |>
  pivot_wider(names_from = year,
              values_from = prop)

prop_mat <- as.matrix(prop_df[, 2:ncol(prop_df)])
rownames(prop_mat) <- prop_df$name

# create heatmap
pheatmap(prop_mat, cluster_cols = FALSE, scale = "row")
```

In @fig-line-2010-1 and @fig-line-2010-1 the line plots are shown. To view trends of many names at once, @fig-line-2010-3 displays a heatmap.
